---
title: "RSV_250131"
---

### load in packages
```{r}
library(arrow)
library(dplyr)
library(survival)
library(survminer)
library(tidyr)
```

### read in parquet database
```{r}
# count: 9359293
NY_pq <- open_dataset("/Volumes/home/HCUPdata-CC0941-MEDSPH/ParquetData/NY", format = "parquet")
NY_pq_df <- as.data.frame(NY_pq)
```

# Check birth year distribution of the infants.
```{r}
NY_pq_df %>%
  filter(`Birth year` %in% c(2016, 2017, 2018, 2019)) %>% 
  mutate(BWT_category = case_when(
    `Birth weight in grams` == "" ~ "Empty",
    `Birth weight in grams` == "A" ~ "A",
    suppressWarnings(!is.na(as.numeric(`Birth weight in grams`))) ~ "Numeric",
    TRUE ~ "Other"
  )) %>%
  group_by(`Birth year`, BWT_category) %>%
  summarise(distinct_visits = n_distinct(`Visit linkage variable`), .groups = "drop") %>% 
  arrange(`Birth year`)

```


```{r}
# count: 840841
NY_cohort1 <- NY_pq_df %>%
  select('Visit linkage variable', 'Birth year', 'Birth month', 'Birth weight in grams') %>%
  rename(
    ID = 'Visit linkage variable',
    birth_year = 'Birth year',
    birth_month = 'Birth month',
    BWT = 'Birth weight in grams'
  ) %>%
  filter(BWT != '' & !is.na(BWT)) %>%
  distinct(ID, .keep_all = TRUE)  
```

```{r}
# factors and details
# count:5375860
NY_cohort2 <- NY_pq_df %>%
  select('Visit linkage variable', 'Birth year', 'Birth month', 'Birth weight in grams',
         'Admission month', 'Admission year', 'Indicator of sex', 'Race (uniform)',
         'Hispanic ethnicity (uniform)', 'Median household income state quartile for patient ZIP Code',
         'Median household income national quartile for patient ZIP Code', 'Patient ZIP Code',
         'Patient ZIP Code, first 3 digits', 'Patient State postal code', 'Died during hospitalization',
         'Calendar year','Discharge month','Age in years at admission','Age in days (when age < 1 year)','Age in months (when age < 11 years)',
         starts_with("Diagnosis"), starts_with("ICD-10-CM Diagnosis")) %>%
  rename(
    # Basic details
    ID = 'Visit linkage variable',
    Amonth = 'Admission month',
    Ayear = 'Admission year',
    birth_year = 'Birth year',
    birth_month = 'Birth month',
    BWT = 'Birth weight in grams',
    Sex = 'Indicator of sex',
    Race = 'Race (uniform)',
    Ethnicity = 'Hispanic ethnicity (uniform)',
    Income_state_quartile = 'Median household income state quartile for patient ZIP Code',
    Income_national_quartile = 'Median household income national quartile for patient ZIP Code',
    ZIP = 'Patient ZIP Code',
    ZIP_3digits = 'Patient ZIP Code, first 3 digits',
    State = 'Patient State postal code',
    Death = 'Died during hospitalization',
    Dyear = 'Calendar year',
    Dmonth = 'Discharge month',
    admis_age_year = 'Age in years at admission',
    admis_age_day = 'Age in days (when age < 1 year)',
    admis_age_mon = 'Age in months (when age < 11 years)',
  ) %>%
  distinct(ID, .keep_all = TRUE) %>%
  collect()
```

```{r}
# Kept only rows with BWT data and distincted
# 840841
NY_merged <- NY_cohort1 %>%
  left_join(NY_cohort2, by = "ID")

saveRDS(NY_merged, '/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_merged_20250131.rds')
```

# # # 
```{r}
NY_merged <- readRDS("/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_merged_20250131.rds")
```


```{r}
NY_merged %>%
  count(birth_year.x) %>%
  filter(birth_year.x %in% c(2014, 2015, 2016, 2017, 2018, 2019)) %>%
  print()

# State number: 
# 2016 - 232,663
# 2017 - 228,501
# 2018 - 225,162
# 2019 - 220,536

# Database number:
# 2016 - 220,343	
# 2017 - 214,468
# 2018 - 194,315	
# 2019 - 192,325	
```
```{r}
NY_merged %>%
  count(Dyear) %>%
  filter(Dyear %in% c(2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021)) %>%
  print()
```

=======

```{r}
# Turn age to numeric and normalize to day unit
NY_cohort3 <- NY_merged %>% 
  mutate(
    admis_age_year = if_else(admis_age_year %in% c("A", "B", "C"), NA_character_, admis_age_year),
    admis_age_mon = if_else(admis_age_mon %in% c("A", "B", "C"), NA_character_, admis_age_mon),
    admis_age_day = if_else(admis_age_day %in% c("A", "B", "C"), NA_character_, admis_age_day)
  ) %>%
  mutate(
    admis_age_year = as.numeric(admis_age_year),
    admis_age_mon = as.numeric(admis_age_mon),
    admis_age_day = as.numeric(admis_age_day)
  )

# make sure only infant <2 years are included
NY_cohort4 <- NY_cohort3 %>% 
  mutate(
    admis_age_year = if_else(admis_age_year > 2, NA_real_, admis_age_year),
    admis_age_mon = if_else(admis_age_mon > 24, NA_real_, admis_age_mon),
    admis_age_day = if_else(admis_age_day > 730, NA_real_, admis_age_day),
    
    admis_age_norm = pmax(
      admis_age_year * 365.25,
      admis_age_mon * 30.44,
      admis_age_day,
      na.rm = TRUE
    )
  ) 

```

### Finds rsv encounters each patient has and flag as 1, count number
```{r}
# Step 1: Rename ICD columns to a simpler format
NY_cohort4 <- NY_cohort4 %>%
  rename_with(~ gsub("ICD-10-CM Diagnosis ", "ICD_", .), starts_with("ICD-10-CM Diagnosis"))
NY_cohort4 <- NY_cohort4 %>%
  rename_with(~ gsub("Diagnosis (\\d+), present on admission indicator", "Diag_admis_\\1", .), everything())
```

```{r}
saveRDS(NY_cohort4, '/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_cohort4_20250210.rds')
```

```{r}
NY_cohort4 <- readRDS("/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_cohort4_20250210.rds")
```

### RSV indi [Problem need to be solved]
```{r}
# RSV indicator as a new column
NY_cohort4 <- NY_cohort4 %>%
  mutate(
    rsv_bronchitis = as.integer(rowSums(across(ICD_1:ICD_25, ~ . == "J205"), na.rm = TRUE) > 0),
    rsv_bronchiolitis = as.integer(rowSums(across(ICD_1:ICD_25, ~ . == "J210"), na.rm = TRUE) > 0),
    rsv_pneumonia = as.integer(rowSums(across(ICD_1:ICD_25, ~ . == "J121"), na.rm = TRUE) > 0),
    rsv_cause_of_disease = as.integer(rowSums(across(ICD_1:ICD_25, ~ . == "B974"), na.rm = TRUE) > 0),
    rsv_indi = as.integer(rsv_bronchitis + rsv_bronchiolitis + rsv_pneumonia + rsv_cause_of_disease > 0)
  )
```

```{r}
NY_cohort4 %>%
  summarise(
    total = n(),
    rsv_bronchitis = sum(rsv_bronchitis, na.rm = TRUE),
    rsv_bronchiolitis = sum(rsv_bronchiolitis, na.rm = TRUE),
    rsv_pneumonia = sum(rsv_pneumonia, na.rm = TRUE),
    rsv_cause_of_disease = sum(rsv_cause_of_disease, na.rm = TRUE),
    rsv_indi = sum(rsv_indi, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = -total, names_to = "RSV_Type", values_to = "Count") %>%
  mutate(Percentage = (Count / total) * 100) %>%
  select(RSV_Type, Count, Percentage) %>%
  print()
```


```{r}
# Step 2: Convert RSV-related ICD codes to binary indicators
NY_cohort5 <- NY_cohort4 %>%
  mutate(across(
    starts_with("ICD_"),
    ~ ifelse(. %in% c("J205", "J210", "J121", "B974"), 1, 0),
    .names = "rsv_{.col}"
  ))

# Exclude columns that are no longer needed
NY_cohort5 <- NY_cohort5 %>%
  select(-all_of(paste0("ICD_", 1:25)))
```


# Total encounter of RSV
NY_cohort5 <- NY_cohort5 %>%
  mutate(rsv_num = rowSums(select(., starts_with("rsv_ICD"))))
NY_cohort5 <- NY_cohort5 %>%
  select(-all_of(paste0("rsv_ICD_", 1:25)))


### Flag rows of patient who were admitted with rsv
```{r}
NY_cohort5 <- NY_cohort5 %>%
  mutate(across(starts_with("Diag_admis_"), ~ as.integer(. == "Y"), .names = "check_{.col}")) 

# Exclude columns that are no longer needed
NY_cohort5 <- NY_cohort5 %>%
  select(-all_of(paste0("Diag_admis_", 1:25)))
```


```{r}
# Flag people got RSV before admission
NY_cohort5 <- NY_cohort5 %>%
  mutate(
    rsv_admis = as.integer(
      rowSums(
        select(., starts_with("rsv_ICD_")) * select(., starts_with("check_Diag_")) > 0,
        na.rm = TRUE
      ) > 0
    )
  )
```

```{r}
# Exclude columns that are no longer needed
NY_cohort6 <- NY_cohort5 %>%
  select(-all_of(paste0("check_Diag_admis_", 1:25)))
NY_cohort6 <- NY_cohort6 %>%
  select(-all_of(paste0("rsv_ICD_", 1:25)))

NY_cohort6 <- NY_cohort6 %>%
  select(-birth_year.y, -birth_month.y) %>% 
  rename(
    birth_year = birth_year.x, 
    birth_month = birth_month.x 
  )

```

# QUESTION: ALL INFANT GOT RSV BEFORE ADMISSION???
count_mismatch <- sum(NY_cohort6$rsv_admis != NY_cohort6$rsv_indic, na.rm = TRUE)
print(count_mismatch)

```{r}
saveRDS(NY_cohort6, '/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_cohort6_20250210.rds')
```

```{r}
NY_cohort6 <- readRDS("/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_cohort6_20250210.rds")
```


### KM Curve
```{r}
# Filter out rows where birth_year < 2014 count:821135
NY_cohort7 <- NY_cohort6 %>% filter(birth_year >= 2014)
```

```{r}
# Create start and end time in YYYY.MM format
NY_cohort7 <- NY_cohort7 %>% 
  mutate(start_time = birth_year + birth_month / 12,
         end_time = ifelse(rsv_admis == 1, Ayear + Amonth / 12, Dyear + Dmonth / 12),
         time_to_event = end_time - start_time)

# Create survival object
surv_obj <- Surv(time = NY_cohort7$time_to_event, event = NY_cohort7$rsv_admis)

# Fit Kaplan-Meier model
km_fit <- survfit(surv_obj ~ 1, data = NY_cohort7)

# Plot KM curve
ggsurvplot(km_fit, 
           data = NY_cohort7, 
           xlab = "Time in Years", 
           ylab = "Survival Probability", 
           title = "Kaplan-Meier Curve for RSV Patients", 
           conf.int = TRUE,
           risk.table = TRUE)  # Ensures the risk table is properly displayed
```
# Variables clean up
```{r}
NY_cohort7 <- NY_cohort7 %>%
  mutate(
    Sex = case_when(
      Sex == "0" ~ "Male",
      Sex == "1" ~ "Female",
      TRUE ~ NA_character_
    ),
    Race = case_when(
      Race == "1" ~ "White",
      Race == "2" ~ "Black",
      Race == "3" ~ "Hispanic",
      Race == "4" ~ "Asian or Pacific Islander",
      Race == "5" ~ "Native American",
      Race == "6" ~ "Other",
      TRUE ~ NA_character_
    ),
    Ethnicity = case_when(
      Ethnicity == 0 ~ "Not Hispanic",
      Ethnicity == 1 ~ "Hispanic, White",
      Ethnicity == 2 ~ "Hispanic, Black",
      Ethnicity == 3 ~ "Hispanic, Other Race",
      Ethnicity == 4 ~ "Hispanic, Unspecified Race",
      TRUE ~ NA_character_ 
    ),
    N_Quartile = case_when(
      Income_national_quartile == 1 ~ "First Quartile",
      Income_national_quartile == 2 ~ "Second Quartile",
      Income_national_quartile == 3 ~ "Third Quartile",
      Income_national_quartile == 4 ~ "Fourth Quartile",
      TRUE ~ NA_character_ 
    ),
    S_Quartile = case_when(
      Income_state_quartile == 1 ~ "First Quartile",
      Income_state_quartile == 2 ~ "Second Quartile",
      Income_state_quartile == 3 ~ "Third Quartile",
      Income_state_quartile == 4 ~ "Fourth Quartile",
      TRUE ~ NA_character_ 
    )
  )
```

```{r}
# Fit the KM model stratified by Sex
km_fit_sex <- survfit(surv_obj ~ Sex, data = NY_cohort7)
# Plot the KM curve stratified by Sex
ggsurvplot(
  km_fit_sex,
  data = NY_cohort7,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Year)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by Sex"
)
```


```{r}
# Fit the KM model stratified by Race
km_fit_race <- survfit(surv_obj ~ Race, data = NY_cohort7)
# Plot the KM curve stratified by Race
ggsurvplot(
  km_fit_race,
  data = NY_cohort7,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Year)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by Race"
)
```

```{r}
# Fit the KM model stratified by Race
km_fit_ethnicity <- survfit(surv_obj ~ Ethnicity, data = NY_cohort7)

# Plot the KM curve stratified by Race
ggsurvplot(
  km_fit_ethnicity,
  data = NY_cohort7,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Year)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by Ethnicity"
)
```

```{r}
# Fit the KM model stratified by Race
km_fit_n_income_Quartile <- survfit(surv_obj ~ N_Quartile, data = NY_cohort7)

# Plot the KM curve stratified by Race
ggsurvplot(
  km_fit_n_income_Quartile,
  data = NY_cohort7,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Year)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by National Income Quartile"
)
```

```{r}
# Fit the KM model stratified by Race
km_fit_s_income_Quartile <- survfit(surv_obj ~ S_Quartile, data = NY_cohort7)

# Plot the KM curve stratified by Race
ggsurvplot(
  km_fit_s_income_Quartile,
  data = NY_cohort7,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Year)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by State Income Quartile"
)
```
