---
title: "RSV_Cohort_Explore_250211"
output: pdf_document
date: "2025-02-11"
---
### load in packages
```{r}
library(arrow)
library(dplyr)
library(survival)
library(survminer)
library(tidyr)
```

```{r}
# count: 9359293
NY_pq <- open_dataset("/Volumes/home/HCUPdata-CC0941-MEDSPH/ParquetData/NY", format = "parquet")
NY_pq_df <- as.data.frame(NY_pq)
```

```{r}
# count: 840841
NY_cohort1 <- NY_pq_df %>%
  select('Visit linkage variable', 'Birth year', 'Birth month', 'Birth weight in grams') %>%
  rename(
    ID = 'Visit linkage variable',
    birth_year = 'Birth year',
    birth_month = 'Birth month',
    BWT = 'Birth weight in grams'
  ) %>%
  filter(BWT != '' & !is.na(BWT)) %>%
  distinct(ID, .keep_all = TRUE)  
```

```{r Mian change: Kept duplicated rolls}
# factors and details
# count:9,359,293
NY_cohort2_2 <- NY_pq_df %>%
  select(`Visit linkage variable`, `Birth year`, `Birth month`, `Birth weight in grams`,
         `Admission month`, `Admission year`, `Indicator of sex`, `Race (uniform)`,
         `Hispanic ethnicity (uniform)`, `Median household income state quartile for patient ZIP Code`,
         `Median household income national quartile for patient ZIP Code`, `Patient ZIP Code`,
         `Patient ZIP Code, first 3 digits`, `Patient State postal code`, `Died during hospitalization`,
         `Calendar year`, `Discharge month`, `Age in years at admission`, `Age in days (when age < 1 year)`,
         `Age in months (when age < 11 years)`, starts_with("Diagnosis"), starts_with("ICD-10-CM Diagnosis")) %>%
  rename(
    # Basic details
    ID = `Visit linkage variable`,
    Amonth = `Admission month`,
    Ayear = `Admission year`,
    birth_year = `Birth year`,
    birth_month = `Birth month`,
    BWT = `Birth weight in grams`,
    Sex = `Indicator of sex`,
    Race = `Race (uniform)`,
    Ethnicity = `Hispanic ethnicity (uniform)`,
    Income_state_quartile = `Median household income state quartile for patient ZIP Code`,
    Income_national_quartile = `Median household income national quartile for patient ZIP Code`,
    ZIP = `Patient ZIP Code`,
    ZIP_3digits = `Patient ZIP Code, first 3 digits`,
    State = `Patient State postal code`,
    Death = `Died during hospitalization`,
    Dyear = `Calendar year`,
    Dmonth = `Discharge month`,
    admis_age_year = `Age in years at admission`,
    admis_age_day = `Age in days (when age < 1 year)`,
    admis_age_mon = `Age in months (when age < 11 years)`
  )
```

# Separate and treat each ICD code separately 
```{r}
# rsv_bronchitis count: 
NY_cohort2_J205 <- NY_cohort2_2 %>%
  unite('all_icd', starts_with('ICD'), sep = ";", remove = FALSE) %>%
  mutate(J205 = if_else(grepl('J205', all_icd), 1, 0)) %>%
  group_by(`ID`) %>%
  slice_max(J205, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(ID, J205)

saveRDS(NY_cohort2_J205, '/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_cohort2_J205_250211')
```

```{r}
# rsv_bronchiolitis count
NY_cohort2_J210 <- NY_cohort2_2 %>%
  unite('all_icd', starts_with('ICD'), sep = ";", remove = FALSE) %>%
  mutate(J210 = if_else(grepl('J210', all_icd), 1, 0)) %>%
  group_by(`ID`) %>%
  slice_max(J210, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(ID, J210)

saveRDS(NY_cohort2_J210, '/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_cohort2_J210_250211')
```

```{r}
# Acute bronchiolitis due to human metapneumovirus count
NY_cohort2_J211 <- NY_cohort2_2 %>%
  unite('all_icd', starts_with('ICD'), sep = ";", remove = FALSE) %>%
  mutate(J211 = if_else(grepl('J211', all_icd), 1, 0)) %>%
  group_by(`ID`) %>%
  slice_max(J211, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(ID, J211)

saveRDS(NY_cohort2_J211, '/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_cohort2_J211_250211')
```

```{r}
# Acute bronchiolitis due to other specified organisms count
NY_cohort2_J218 <- NY_cohort2_2 %>%
  unite('all_icd', starts_with('ICD'), sep = ";", remove = FALSE) %>%
  mutate(J218 = if_else(grepl('J218', all_icd), 1, 0)) %>%
  group_by(`ID`) %>%
  slice_max(J218, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(ID, J218)

saveRDS(NY_cohort2_J218, '/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_cohort2_J218_250211')
```

```{r}
# Acute bronchiolitis, unspecified count
NY_cohort2_J219 <- NY_cohort2_2 %>%
  unite('all_icd', starts_with('ICD'), sep = ";", remove = FALSE) %>%
  mutate(J219 = if_else(grepl('J219', all_icd), 1, 0)) %>%
  group_by(`ID`) %>%
  slice_max(J219, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(ID, J219)

saveRDS(NY_cohort2_J219, '/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_cohort2_J219_250211')
```

```{r}
# rsv_pneumonia count
NY_cohort2_J121 <- NY_cohort2_2 %>%
  unite('all_icd', starts_with('ICD'), sep = ";", remove = FALSE) %>%
  mutate(J121 = if_else(grepl('J121', all_icd), 1, 0)) %>%
  group_by(`ID`) %>%
  slice_max(J121, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(ID, J121)

saveRDS(NY_cohort2_J121, '/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_cohort2_J121_250211')
```

```{r}
# rsv_cause_of_disease count
NY_cohort2_B974 <- NY_cohort2_2 %>%
  unite('all_icd', starts_with('ICD'), sep = ";", remove = FALSE) %>%
  mutate(B974 = if_else(grepl('B974', all_icd), 1, 0)) %>%
  group_by(`ID`) %>%
  slice_max(B974, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(ID, B974)

saveRDS(NY_cohort2_B974, '/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_cohort2_B974_250211')
```

# load in code for future use
NY_cohort2_J205 <- readRDS('/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_cohort2_J205_250211')
NY_cohort2_J210 <- readRDS('/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_cohort2_J210_250211')
NY_cohort2_J211 <- readRDS('/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_cohort2_J211_250211')
NY_cohort2_J218 <- readRDS('/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_cohort2_J218_250211')
NY_cohort2_J219 <- readRDS('/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_cohort2_J219_250211')
NY_cohort2_J121 <- readRDS('/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_cohort2_J121_250211')
NY_cohort2_B974 <- readRDS('/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_cohort2_B974_250211')

# merge, limit id to infant <2ys
```{r}
NY_cohort2_3 <- NY_cohort1 %>%
  left_join(NY_cohort2_B974, by = "ID") %>%
  left_join(NY_cohort2_J121, by = "ID") %>%
  left_join(NY_cohort2_J205, by = "ID") %>%
  left_join(NY_cohort2_J210, by = "ID") %>%
  left_join(NY_cohort2_J211, by = "ID") %>%
  left_join(NY_cohort2_J218, by = "ID") %>%
  left_join(NY_cohort2_J219, by = "ID")

NY_cohort2_3 <- NY_cohort2_3 %>%
  mutate(
    rsv_indi = as.integer(B974 + J121 + J205 + J210 + J211 + J218 + J219 > 0)
  )
```

# summary
```{r}
count_summary <- NY_cohort2_3 %>%
  summarise(across(c(B974, J121, J205, J210, J211, J218, J219, rsv_indi),
                   list(count_0 = ~sum(. == 0, na.rm = TRUE),
                        count_1 = ~sum(. == 1, na.rm = TRUE),
                        pct_0 = ~mean(. == 0, na.rm = TRUE) * 100,
                        pct_1 = ~mean(. == 1, na.rm = TRUE) * 100))) %>%
  pivot_longer(cols = everything(), 
               names_to = c("Variable", ".value"),
               names_pattern = "(.*)_(count_0|count_1|pct_0|pct_1)")

# Display table in console
print(count_summary)

```

# Distinct ID
```{r}
NY_cohort2_4 <- NY_cohort2_2 %>%
  distinct(ID, .keep_all = TRUE)
```
# Merge
```{r}
NY_cohort2_5 <- NY_cohort2_3 %>%
  left_join(NY_cohort2_4, by = "ID")

NY_cohort2_5 <- NY_cohort2_5 %>%
  select(ID, birth_year.x, birth_month.x, BWT.x, B974, J121, J205, J210, J211, J218, J219, rsv_indi, 
         birth_year.y, birth_month.y, BWT.y, Amonth, Ayear, Sex, Race, Ethnicity, 
         Income_state_quartile, Income_national_quartile, ZIP, ZIP_3digits, 
         State, Death, Dyear, Dmonth, admis_age_year, admis_age_day, admis_age_mon)

```

# Normalize age.
```{r}
# Turn age to numeric and normalize to day unit
NY_cohort3 <- NY_cohort2_5 %>% 
  mutate(
    admis_age_year = if_else(admis_age_year %in% c("A", "B", "C"), NA_character_, admis_age_year),
    admis_age_mon = if_else(admis_age_mon %in% c("A", "B", "C"), NA_character_, admis_age_mon),
    admis_age_day = if_else(admis_age_day %in% c("A", "B", "C"), NA_character_, admis_age_day)
  ) %>%
  mutate(
    admis_age_year = as.numeric(admis_age_year),
    admis_age_mon = as.numeric(admis_age_mon),
    admis_age_day = as.numeric(admis_age_day)
  )

# make sure only infant <2 years are included
NY_cohort4 <- NY_cohort3 %>% 
  mutate(
    admis_age_year = if_else(admis_age_year > 2, NA_real_, admis_age_year),
    admis_age_mon = if_else(admis_age_mon > 24, NA_real_, admis_age_mon),
    admis_age_day = if_else(admis_age_day > 730, NA_real_, admis_age_day),
    
    admis_age_norm = pmax(
      admis_age_year * 365.25,
      admis_age_mon * 30.44,
      admis_age_day,
      na.rm = TRUE
    )
  ) %>%
  filter(birth_year.x >= 2014 & birth_year.x <= 2019)

# Exclude unewanted rows
NY_cohort4 <- NY_cohort4 %>%
  select(-birth_year.y, -birth_month.y,-BWT.y) %>% 
  rename(
    birth_year = birth_year.x, 
    birth_month = birth_month.x, 
    BWT = BWT.x
  )
```


# KM Curve
```{r}
# Create start and end time in YYYY.MM format
NY_cohort4 <- NY_cohort4 %>% 
  mutate(start_time = birth_year + birth_month / 12,
         end_time = ifelse(rsv_indi == 1, Ayear + Amonth / 12, Dyear + Dmonth / 12),
         time_to_event = end_time - start_time)
```

```{r}
saveRDS(NY_cohort4, '/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_cohort4_20250211.rds')
```

```{r}
NY_cohort4 <- readRDS("/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/NY_cohort4_20250211.rds")
```

```{r}
# Create survival object
surv_obj <- Surv(time = NY_cohort4$time_to_event, event = NY_cohort4$rsv_indi)

# Fit Kaplan-Meier model
km_fit <- survfit(surv_obj ~ 1, data = NY_cohort4)

# Plot KM curve
ggsurvplot(km_fit, 
           data = NY_cohort4, 
           xlab = "Time in Years", 
           ylab = "Survival Probability", 
           title = "Kaplan-Meier Curve for RSV Patients", 
           conf.int = TRUE,
           risk.table = TRUE)  # Ensures the risk table is properly displayed
```
           
# Variables clean up
```{r}
NY_cohort4 <- NY_cohort4 %>%
  mutate(
    Sex = case_when(
      Sex == "0" ~ "Male",
      Sex == "1" ~ "Female",
      TRUE ~ NA_character_
    ),
    Race = case_when(
      Race == "1" ~ "White",
      Race == "2" ~ "Black",
      Race == "3" ~ "Hispanic",
      Race == "4" ~ "Asian or Pacific Islander",
      Race == "5" ~ "Native American",
      Race == "6" ~ "Other",
      TRUE ~ NA_character_
    ),
    Ethnicity = case_when(
      Ethnicity == 0 ~ "Not Hispanic",
      Ethnicity == 1 ~ "Hispanic, White",
      Ethnicity == 2 ~ "Hispanic, Black",
      Ethnicity == 3 ~ "Hispanic, Other Race",
      Ethnicity == 4 ~ "Hispanic, Unspecified Race",
      TRUE ~ NA_character_ 
    ),
    N_Quartile = case_when(
      Income_national_quartile == 1 ~ "First Quartile",
      Income_national_quartile == 2 ~ "Second Quartile",
      Income_national_quartile == 3 ~ "Third Quartile",
      Income_national_quartile == 4 ~ "Fourth Quartile",
      TRUE ~ NA_character_ 
    ),
    S_Quartile = case_when(
      Income_state_quartile == 1 ~ "First Quartile",
      Income_state_quartile == 2 ~ "Second Quartile",
      Income_state_quartile == 3 ~ "Third Quartile",
      Income_state_quartile == 4 ~ "Fourth Quartile",
      TRUE ~ NA_character_ 
    )
  )
```

```{r}
# Fit the KM model stratified by Sex
km_fit_sex <- survfit(surv_obj ~ Sex, data = NY_cohort4)
# Plot the KM curve stratified by Sex
ggsurvplot(
  km_fit_sex,
  data = NY_cohort4,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Year)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by Sex"
)
```

```{r}
NY_cohort4 <- NY_cohort4 %>%
  mutate(Race = factor(Race, 
                       levels = c("White", "Black", "Asian or Pacific Islander", 
                                  "Hispanic", "Native American", "Other"),
                       ordered = TRUE))
# Fit the KM model stratified by Race
km_fit_race <- survfit(surv_obj ~ Race, data = NY_cohort4)
# Plot the KM curve stratified by Race
ggsurvplot(
  km_fit_race,
  data = NY_cohort4,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Year)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by Race"
)
```

```{r}
NY_cohort4 <- NY_cohort4 %>%
  mutate(Ethnicity = factor(Ethnicity, 
                            levels = c("Hispanic, White",
                                       "Hispanic, Black", 
                                       "Hispanic, Other Race", 
                                       "Not Hispanic"),
                            ordered = TRUE))

# Fit the KM model stratified by Race
km_fit_ethnicity <- survfit(surv_obj ~ Ethnicity, data = NY_cohort4)

# Plot the KM curve stratified by Race
ggsurvplot(
  km_fit_ethnicity,
  data = NY_cohort4,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Year)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by Ethnicity"
)
```

```{r}
NY_cohort4 <- NY_cohort4 %>%
  mutate(N_Quartile = factor(N_Quartile, 
                             levels = c("First Quartile", "Second Quartile", "Third Quartile", "Fourth Quartile"), 
                             ordered = TRUE))

# Fit the KM model stratified by Race
km_fit_n_income_Quartile <- survfit(surv_obj ~ N_Quartile, data = NY_cohort4)

# Plot the KM curve stratified by Race
ggsurvplot(
  km_fit_n_income_Quartile,
  data = NY_cohort4,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Year)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by National Income Quartile"
)
```

```{r}
NY_cohort4 <- NY_cohort4 %>%
  mutate(S_Quartile = factor(S_Quartile, 
                             levels = c("First Quartile", "Second Quartile", "Third Quartile", "Fourth Quartile"), 
                             ordered = TRUE))

# Fit the KM model stratified by Race
km_fit_s_income_Quartile <- survfit(surv_obj ~ S_Quartile, data = NY_cohort4)

# Plot the KM curve stratified by Race
ggsurvplot(
  km_fit_s_income_Quartile,
  data = NY_cohort4,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Year)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by State Income Quartile"
)
```

# Start from here only see patients that got a NY zip code 100-104; 110-119
```{r}
NY_cohort5 <- NY_cohort4 %>%
  filter(ZIP_3digits %in% c(100:104, 110:119))
```

# KM Curve
```{r}
# Create survival object
surv_obj <- Surv(time = NY_cohort5$time_to_event, event = NY_cohort5$rsv_indi)

# Fit Kaplan-Meier model
km_fit <- survfit(surv_obj ~ 1, data = NY_cohort5)

# Plot KM curve
ggsurvplot(km_fit, 
           data = NY_cohort5, 
           xlab = "Time in Years", 
           ylab = "Survival Probability", 
           title = "Kaplan-Meier Curve for RSV Patients", 
           conf.int = TRUE,
           risk.table = TRUE) 
```

# Variables clean up
```{r}
NY_cohort5 <- NY_cohort5 %>%
  mutate(
    Sex = case_when(
      Sex == "0" ~ "Male",
      Sex == "1" ~ "Female",
      TRUE ~ NA_character_
    ),
    Race = case_when(
      Race == "1" ~ "White",
      Race == "2" ~ "Black",
      Race == "3" ~ "Hispanic",
      Race == "4" ~ "Asian or Pacific Islander",
      Race == "5" ~ "Native American",
      Race == "6" ~ "Other",
      TRUE ~ NA_character_
    ),
    Ethnicity = case_when(
      Ethnicity == 0 ~ "Not Hispanic",
      Ethnicity == 1 ~ "Hispanic, White",
      Ethnicity == 2 ~ "Hispanic, Black",
      Ethnicity == 3 ~ "Hispanic, Other Race",
      Ethnicity == 4 ~ "Hispanic, Unspecified Race",
      TRUE ~ NA_character_ 
    ),
    N_Quartile = case_when(
      Income_national_quartile == 1 ~ "First Quartile",
      Income_national_quartile == 2 ~ "Second Quartile",
      Income_national_quartile == 3 ~ "Third Quartile",
      Income_national_quartile == 4 ~ "Fourth Quartile",
      TRUE ~ NA_character_ 
    ),
    S_Quartile = case_when(
      Income_state_quartile == 1 ~ "First Quartile",
      Income_state_quartile == 2 ~ "Second Quartile",
      Income_state_quartile == 3 ~ "Third Quartile",
      Income_state_quartile == 4 ~ "Fourth Quartile",
      TRUE ~ NA_character_ 
    )
  )
```

```{r}
# Fit the KM model stratified by Sex
km_fit_sex <- survfit(surv_obj ~ Sex, data = NY_cohort5)
# Plot the KM curve stratified by Sex
ggsurvplot(
  km_fit_sex,
  data = NY_cohort5,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Year)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by Sex"
)
```

```{r}
NY_cohort5 <- NY_cohort5 %>%
  mutate(Race = factor(Race, 
                       levels = c("White", "Black", "Asian or Pacific Islander", 
                                  "Hispanic", "Native American", "Other"),
                       ordered = TRUE))

# Fit the KM model stratified by Race
km_fit_race <- survfit(surv_obj ~ Race, data = NY_cohort5)
# Plot the KM curve stratified by Race
ggsurvplot(
  km_fit_race,
  data = NY_cohort5,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Year)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by Race"
)
```

```{r}
NY_cohort5 <- NY_cohort5 %>%
  mutate(Ethnicity = factor(Ethnicity, 
                            levels = c("Hispanic, White",
                                       "Hispanic, Black", 
                                       "Hispanic, Other Race", 
                                       "Not Hispanic"),
                            ordered = TRUE))

# Fit the KM model stratified by Race
km_fit_ethnicity <- survfit(surv_obj ~ Ethnicity, data = NY_cohort5)

# Plot the KM curve stratified by Race
ggsurvplot(
  km_fit_ethnicity,
  data = NY_cohort5,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Year)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by Ethnicity"
)
```

```{r}
NY_cohort5 <- NY_cohort5 %>%
  mutate(N_Quartile = factor(N_Quartile, 
                             levels = c("First Quartile", "Second Quartile", "Third Quartile", "Fourth Quartile"), 
                             ordered = TRUE))

# Fit the KM model stratified by Race
km_fit_n_income_Quartile <- survfit(surv_obj ~ N_Quartile, data = NY_cohort5)

# Plot the KM curve stratified by Race
ggsurvplot(
  km_fit_n_income_Quartile,
  data = NY_cohort5,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Year)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by National Income Quartile"
)
```

```{r}
NY_cohort5 <- NY_cohort5 %>%
  mutate(S_Quartile = factor(S_Quartile, 
                             levels = c("First Quartile", "Second Quartile", "Third Quartile", "Fourth Quartile"), 
                             ordered = TRUE))

# Fit the KM model stratified by Race
km_fit_s_income_Quartile <- survfit(surv_obj ~ S_Quartile, data = NY_cohort5)

# Plot the KM curve stratified by Race
ggsurvplot(
  km_fit_s_income_Quartile,
  data = NY_cohort5,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Year)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by State Income Quartile"
)
```

NY_cohort5 <- NY_cohort5 %>%
  mutate(S_Quartile = factor(S_Quartile, 
                             levels = c("First Quartile", "Second Quartile", "Third Quartile", "Fourth Quartile"), 
                             ordered = TRUE))

# Fit the KM model stratified by Race
km_fit_s_income_Quartile <- survfit(surv_obj ~ S_Quartile, data = NY_cohort5)

# Plot the KM curve stratified by Race
ggsurvplot(
  km_fit_s_income_Quartile,
  data = NY_cohort5,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Year)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by State Income Quartile"
)
