---
title: "MergedTable4" - First step
output: html_document
date: "2025-03-06"
---
```{r}
library(arrow)
library(dplyr)
library(survival)
library(survminer)
library(tidyr)
library(stringr)
```

# SID parquet
```{r}
# # count: 9359293
# NY_pq <- open_dataset("/Volumes/home/HCUPdata-CC0941-MEDSPH/ParquetData/NY", format = "parquet")
```

# birth cohort, query from SID table
```{r}
# NY_cohort1 <- NY_pq %>%
#   select('Visit linkage variable', 'Birth year', 'Birth month', 'Birth weight in grams') %>%
#   rename(
#     ID = 'Visit linkage variable',
#     birth_year = 'Birth year',
#     birth_month = 'Birth month',
#     BWT = 'Birth weight in grams'
#   ) %>%
#   filter(BWT != '' & !is.na(BWT) & !is.na(ID)) %>%
#   collect() %>%
#   distinct(ID, .keep_all = TRUE) %>%
#   collect()
# 
# # lower case for easy join
# NY_cohort1 <- NY_cohort1 %>%
#   rename_with(tolower)
```

# SID Total pop
```{r}
# # Convert column names to lowercase before processing
# NY_pq <- NY_pq %>%
#   rename_with(tolower)
# 
# # Select relevant columns (ensuring all column names are lowercase)
# SID_rsv_cases <- NY_pq %>%
#   select(
#     `visit linkage variable`, `birth year`, `birth month`, `admission month`, 
#     `admission year`, `indicator of sex`, `race (uniform)`, `hispanic ethnicity (uniform)`, 
#     `median household income national quartile for patient zip code`, 
#     `patient zip code, first 3 digits`, `age in years at admission`, `patient zip code`,
#     `age in months (when age < 11 years)`, `age in days (when age < 1 year)`, 
#     starts_with("diagnosis"), starts_with("icd-10-cm diagnosis")  
#   ) %>%
#   collect() %>%  
#   rename(
#     # Basic details (using lowercase column names after renaming)
#     id = `visit linkage variable`,
#     amonth = `admission month`,
#     ayear = `admission year`,
#     birth_year = `birth year`,
#     birth_month = `birth month`,
#     sex = `indicator of sex`,
#     race = `race (uniform)`,
#     ethnicity = `hispanic ethnicity (uniform)`,
#     income_national_quartile = `median household income national quartile for patient zip code`,
#     zip = `patient zip code`,
#     zip_3digits = `patient zip code, first 3 digits`,
#     admis_age_year = `age in years at admission`,
#     admis_age_day = `age in days (when age < 1 year)`,
#     admis_age_mon = `age in months (when age < 11 years)`
#   )%>%
#   # Ensure `Ayear` and `Amonth` are numeric
#   mutate(across(c(ayear, amonth), as.numeric)) %>%
#   mutate(
#     date = ayear + amonth / 12,  
#     # Create `all_icd` by pasting all `I10_DX` columns together
#     all_icd = apply(select(., starts_with("icd-10-cm diagnosis ")), 1, function(x) paste(na.omit(x), collapse = " "))
#   ) %>%
#   mutate(
#     J205 = if_else(grepl("J205", all_icd), 1, 0),
#     J210 = if_else(grepl("J210", all_icd), 1, 0),
#     J211 = if_else(grepl("J211", all_icd), 1, 0),
#     J218 = if_else(grepl("J218", all_icd), 1, 0),
#     J219 = if_else(grepl("J219", all_icd), 1, 0),
#     J121 = if_else(grepl("J121", all_icd), 1, 0),
#     B974 = if_else(grepl("B974", all_icd), 1, 0),
#     rsv_flag = if_else(J205 == 1 | J210 == 1 | J211 == 1 | J218 == 1 | J219 == 1 | J121 == 1 | B974 == 1, 1, 0)
#   ) %>%
#   right_join(NY_cohort1, by = "id") %>%
#   arrange(id, date) %>%
#   group_by(id) %>%
#   mutate(
#     rsv_flag = replace_na(rsv_flag, 0),
#     rsv_episode_N = cumsum(rsv_flag)
#   ) %>%
#   filter(rsv_episode_N <= 1)
```


# SEDD parquet
```{r}
#NY_pq <- open_dataset("/Volumes/home/HCUPdata-CC0941-MEDSPH/ParquetData/NY_SEDD", format = "parquet")
```

# SEDD Total pop, Give it about 20 min to run
```{r}
# # Convert column names to lowercase before processing
# NY_pq <- NY_pq %>%
#   rename_with(tolower)
# 
# # Select relevant columns (ensuring all column names are lowercase)
# SEDD_rsv_cases <- NY_pq %>%
#   select(
#     visitlink, byear, bmonth, amonth, ayear, female, race, hispanic,
#     zipinc_qrtl, zip, zip3, age, agemonth, ageday,
#     starts_with("i10_dx")
#   ) %>%
#   collect() %>%  # Pull data into R
#   rename(
#     # Basic details (using lowercase column names after renaming)
#     id = visitlink,
#     birth_year = byear,
#     birth_month = bmonth,
#     sex = female,
#     race = race,
#     ethnicity = hispanic,
#     income_national_quartile = zipinc_qrtl,
#     zip = zip,
#     zip_3digits = zip3,
#     admis_age_year = age,
#     admis_age_day = ageday,
#     admis_age_mon = agemonth
#   )%>%
#   # Ensure `Ayear` and `Amonth` are numeric
#   mutate(across(c(ayear, amonth), as.numeric)) %>%
#   mutate(
#     date = ayear + amonth / 12,
#     # Create `all_icd` by pasting all `I10_DX` columns together
#     all_icd = apply(select(., starts_with("i10_dx")), 1, function(x) paste(na.omit(x), collapse = " "))
#   ) %>%
#   mutate(
#     J205 = if_else(grepl("J205", all_icd), 1, 0),
#     J210 = if_else(grepl("J210", all_icd), 1, 0),
#     J211 = if_else(grepl("J211", all_icd), 1, 0),
#     J218 = if_else(grepl("J218", all_icd), 1, 0),
#     J219 = if_else(grepl("J219", all_icd), 1, 0),
#     J121 = if_else(grepl("J121", all_icd), 1, 0),
#     B974 = if_else(grepl("B974", all_icd), 1, 0),
#     rsv_flag = if_else(J205 == 1 | J210 == 1 | J211 == 1 | J218 == 1 | J219 == 1 | J121 == 1 | B974 == 1, 1, 0)
#   ) %>%
#   right_join(NY_cohort1, by = "id") %>%
#   arrange(id, date) %>%
#   group_by(id) %>%
#   mutate(
#     rsv_flag = replace_na(rsv_flag, 0),
#     rsv_episode_N = cumsum(rsv_flag)
#   ) %>%
#   filter(rsv_episode_N <= 1)
```

=================

# SHORT CUT 1
```{r}
#saveRDS(SID_rsv_cases, '/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/SID_rsv_cases_20250412')
SID_rsv_cases <- readRDS("/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/SID_rsv_cases_20250412")
#saveRDS(SEDD_rsv_cases, '/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/SEDD_rsv_cases_20250419')
SEDD_rsv_cases <- readRDS("/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/SEDD_rsv_cases_20250419")
```

=================

# SID keep only infants
```{r}
SID_rsv_filtered <- SID_rsv_cases %>%
  select(id, birth_year.x, birth_month.x, amonth, ayear, sex, race, ethnicity,
         income_national_quartile, zip, zip_3digits, admis_age_year, admis_age_mon,
         admis_age_day, date, rsv_flag, bwt) %>%
  arrange(id, date) %>%
  group_by(id) %>%
  filter(rsv_flag == 1 | row_number() == 1) %>% 
  slice_head(n = 1) %>%   
  ungroup() %>%
  filter(birth_year.x >= 2016 & birth_year.x <= 2019) %>%
  mutate(
    bwt = as.character(bwt),
    Birthweight = case_when(
      bwt == "A" ~ "Unknown",
      !is.na(bwt) & as.numeric(bwt) >= 2500 ~ "Normal",
      !is.na(bwt) & as.numeric(bwt) < 1500 ~ "Very Low Birthweight",
      !is.na(bwt) & as.numeric(bwt) >= 1500 & as.numeric(bwt) < 2500 ~ "Low Birthweight",
      TRUE ~ NA_character_
    )
  )%>%
rename(
  birth_year = birth_year.x,
  birth_month = birth_month.x
)
```
# SID Descriptive
```{r}
library(dplyr)
library(purrr)

SID_rsv_filtered <- SID_rsv_filtered %>%
  mutate(across(-Birthweight, as.factor))

SID_rsv_filtered %>%
  select(where(is.factor), where(is.character)) %>%
  imap(~ {
    cat("\n\n============================\n")
    cat("Variable:", .y, "\n")
    cat("============================\n")
    as.data.frame(table(.x, useNA = "ifany")) %>%
      rename(Value = ".x", Count = "Freq") %>%
      print()
  })
```

# SEDD keep only infants
```{r}
# Filter, group, and keep only the first RSV-positive case
SEDD_rsv_filtered <- SEDD_rsv_cases %>%
  select(
    id, birth_year.y, birth_month.y, amonth, ayear, sex, race, ethnicity,
    income_national_quartile, zip, zip_3digits, admis_age_year,
    admis_age_mon, admis_age_day, date, rsv_flag, bwt
  ) %>%
  arrange(id, date) %>%
  group_by(id) %>%
  slice_max(order_by = rsv_flag, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  filter(birth_year.y >= 2016 & birth_year.y <= 2019) %>%
  mutate(
    bwt = as.character(bwt),
    Birthweight = case_when(
      bwt == "A" ~ "Unknown",
      !is.na(bwt) & as.numeric(bwt) >= 2500 ~ "Normal",
      !is.na(bwt) & as.numeric(bwt) < 1500 ~ "Very Low Birthweight",
      !is.na(bwt) & as.numeric(bwt) >= 1500 & as.numeric(bwt) < 2500 ~ "Low Birthweight",
      TRUE ~ NA_character_
    )
  )%>%
rename(
  birth_year = birth_year.y,
  birth_month = birth_month.y
)

#### UNFINISHED DO FILL UP
```

# SEDD Descriptive
```{r}
library(dplyr)
library(purrr)

SEDD_rsv_filtered <- SEDD_rsv_filtered %>%
  mutate(across(-Birthweight, as.factor))

SEDD_rsv_filtered %>%
  select(where(is.factor), where(is.character)) %>%
  imap(~ {
    cat("\n\n============================\n")
    cat("Variable:", .y, "\n")
    cat("============================\n")
    as.data.frame(table(.x, useNA = "ifany")) %>%
      rename(Value = ".x", Count = "Freq") %>%
      print()
  })
```


# check birth year
birth_year_counts <- SEDD_rsv_filtered %>%
  group_by(birth_year.y) %>%
  summarise(distinct_id_count = n_distinct(id))
  
# Fill the demo info from SID to SEDD = WRONG
```{r}
library(dplyr)
library(stringr)

# Step 1: Join SID and SEDD by 'id', keeping suffixes to distinguish source
merged_data <- SID_rsv_filtered %>%
  left_join(SEDD_rsv_filtered, by = "id", suffix = c(".sid", ".sedd"))

# Step 2: Identify columns to fill (exclude rsv_flag and birth_year/birth_month)
cols_to_fill <- names(merged_data)
cols_to_fill <- cols_to_fill[
  grepl("\\.sedd$", cols_to_fill) &
  !grepl("rsv_flag", cols_to_fill) &
  !grepl("birth_year|birth_month", cols_to_fill)
]

# Step 3: Fill missing values for general columns (use SEDD first, fallback to SID)
merged_data_filled <- merged_data %>%
  mutate(across(
    all_of(cols_to_fill),
    ~ coalesce(.x, get(str_replace(cur_column(), ".sedd", ".sid"))),
    .names = "{str_replace(.col, '.sedd', '')}"
  )) %>%
  
  # Step 4: Special handling for birth_year and birth_month
  mutate(
    birth_year = if_else(is.na(birth_year.sedd), birth_year.sid, birth_year.sedd),
    birth_month = if_else(is.na(birth_month.sedd), birth_month.sid, birth_month.sedd)
  )

# Step 5: Select final columns (cleaned names)
SEDD_rsv_filtered_final <- merged_data_filled %>%
  select(
    id,
    all_of(str_replace(cols_to_fill, ".sedd", "")),  # cleaned general columns
    birth_year,  # specially handled
    birth_month, # specially handled
    rsv_flag = rsv_flag.sedd  # keep RSV flag from SEDD
  )


```


##### Go to PartB #####

# Level source == PASS
```{r}
# SID_rsv_filtered_1 <- SID_rsv_filtered_1 %>%
#   mutate(source = "SID")
# SEDD_rsv_filtered_1 <- SEDD_rsv_filtered_1 %>%
#   mutate(source = "SEDD")
```

                 
# Total cohort
```{r} 
rsv_merged_1 <- bind_rows(SID_rsv_filtered_1, SEDD_rsv_filtered_1)

rsv_merged_1_filtered <- rsv_merged_1 %>%
  arrange(id, date) %>%
  group_by(id) %>%
  filter(
    if (any(rsv_flag == "RSV positive")) {
      row_number() == min(which(rsv_flag == "RSV positive"))
    } else {
      row_number() == 1
    }
  ) %>%
  ungroup()

rsv_merged_1_filtered %>%
  count(rsv_flag)
# Bacl to Part B
```

# Start from here is date setup for KM curves===================================
# Normalize age.
```{r}
# Turn age to numeric and normalize to day unit
rsv_merged_cases_4_1 <- rsv_merged_1_filtered %>%
  mutate(
    admis_age_year = as.numeric(if_else(admis_age_year %in% c("A", "B", "C"), NA_character_, as.character(admis_age_year))),
    admis_age_mon  = as.numeric(if_else(admis_age_mon  %in% c("A", "B", "C"), NA_character_, as.character(admis_age_mon))),
    admis_age_day  = as.numeric(if_else(admis_age_day %in% c("A", "B", "C"), NA_character_, as.character(admis_age_day)))
  )


# make sure only infant <2 years are included
rsv_merged_cases_4_2 <- rsv_merged_cases_4_1 %>% 
  mutate(
    admis_age_year = if_else(admis_age_year > 2, NA_real_, admis_age_year),
    admis_age_mon = if_else(admis_age_mon > 24, NA_real_, admis_age_mon),
    admis_age_day = if_else(admis_age_day > 730, NA_real_, admis_age_day),
    
    admis_age_norm = pmax(
      admis_age_year * 365.25,
      admis_age_mon * 30.44,
      admis_age_day,
      na.rm = TRUE
    )
  ) 
```



# Set up time to event 
```{r}
rsv_merged_cases_4_2 <- rsv_merged_cases_4_2 %>%
  mutate(
    rsv_flag = case_when(
      rsv_flag == "RSV negative" ~ 0,
      rsv_flag == "RSV positive" ~ 1,
      TRUE ~ NA_real_
    )
  )

# Create start and end time 
rsv_merged_cases_4_2 <- rsv_merged_cases_4_2 %>%
  mutate(
    start_time = birth_year + birth_month / 12,
    end_time = ifelse(rsv_flag == 1,
                      ayear + amonth / 12,
                      2019 + 12 / 12),  # Sets December 31, 2019 as end_time
    time_to_event = end_time - start_time,
    time_to_event_days = time_to_event * 365.25  # Convert years to days
  ) 
```

# Time to event f
```{r}
rsv_merged_cases_4_2 <- rsv_merged_cases_4_2 %>%
  mutate(
    time_to_event_days = round(time_to_event_days, 2),
    time_to_event_f = case_when(
      rsv_flag == 1 & !is.na(admis_age_norm) & admis_age_norm != 0 ~ admis_age_norm,
      rsv_flag == 1 & (admis_age_norm == 0 | is.na(admis_age_norm)) ~ time_to_event_days,
      rsv_flag == 0 ~ time_to_event_days
    )
  )
```


# KM curve visualization
```{r}
# Create survival object
surv_obj <- Surv(time = rsv_merged_cases_4_2$time_to_event_f, event = rsv_merged_cases_4_2$rsv_flag)

# Fit Kaplan-Meier model
km_fit <- survfit(surv_obj ~ 1, data = rsv_merged_cases_4_2)

# Plot KM curve
ggsurvplot(km_fit, 
           data = rsv_merged_cases_4_2, 
           xlab = "Time (Day)", 
           ylab = "Survival Probability", 
           title = "Kaplan-Meier Curve for RSV Patients", 
           conf.int = TRUE,
           risk.table = TRUE,
           xlim = c(0, 730),   
           ylim = c(0.92, 1), 
           break.y.by = 0.01 
           )  
```





```{r}
# Fit the KM model stratified by Sex
km_fit_sex <- survfit(surv_obj ~ sex, data = rsv_merged_cases_4_2)

# Plot the KM curve stratified by Sex
ggsurvplot(
  km_fit_sex,
  data = rsv_merged_cases_4_2,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Day)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by Sex",
  xlim = c(0, 730),   
  ylim = c(0.92, 1), 
  break.y.by = 0.01  
)
```

```{r}
rsv_merged_cases_4_2 <- rsv_merged_cases_4_2 %>%
  mutate(Race = factor(Race, 
                       levels = c("White", "Black", "Asian or Pacific Islander", 
                                  "Hispanic", "Native American", "Other"),
                       ordered = TRUE))
# Fit the KM model stratified by Race
km_fit_race <- survfit(surv_obj ~ Race, data = rsv_merged_cases_4_2)
# Plot the KM curve stratified by Race
ggsurvplot(
  km_fit_race,
  data = rsv_merged_cases_4_2,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Day)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by Race",
  xlim = c(0, 730),   
  ylim = c(0.92, 1), 
  break.y.by = 0.01 
)
```

```{r}
rsv_merged_cases_4_2 <- rsv_merged_cases_4_2 %>%
  mutate(Birthweight = factor(Birthweight, 
                       levels = c("Normal", "Low Birthweight", "Very Low Birthweight","Unknown"),
                       ordered = TRUE))

# Fit the KM model stratified by Sex
km_fit_birthweight <- survfit(surv_obj ~ Birthweight, data = rsv_merged_cases_4_2)

# Plot the KM curve stratified by Sex
ggsurvplot(
  km_fit_birthweight,
  data = rsv_merged_cases_4_2,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Day)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by Birth Weight",
  xlim = c(0, 730),   
  ylim = c(0.92, 1), 
  break.y.by = 0.01  
)
```





```{r}
# Recode National Income Quartile as an ordered factor
rsv_merged_cases_4_2 <- rsv_merged_cases_4_2 %>%
  mutate(
    StateIncomeQuartile = factor(
      StateIncomeQuartile,
      levels = c("$1 – $47,999", "$48,000 – $60,999", "$61,000 – $81,999", "$82,000 and above","Missing"),
      ordered = TRUE
    )
  )

# Fit the KM model stratified by National Income Quartile
km_fit_income <- survfit(surv_obj ~ StateIncomeQuartile, data = rsv_merged_cases_4_2)

# Plot the KM curve with custom legend labels
ggsurvplot(
  km_fit_income,
  data = rsv_merged_cases_4_2,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Day)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by National Income Quartile",
  legend.labs = c(
    "Income = $1 – $47,999",
    "Income = $48,000 – $60,999",
    "Income = $61,000 – $81,999",
    "Income = $82,000 and above",
    "Income = Missing"
  ),
  xlim = c(0, 730),
  ylim = c(0.92, 1),
  break.y.by = 0.01
)

```




# Start from here only see patients that got a NY zip code =====
```{r}
rsv_merged_cases_5 <- rsv_merged_cases_4_2 %>% 
  filter(State.Abbreviation == "NY")
```

# KM Curve
```{r}
# Create survival object
surv_obj <- Surv(time = rsv_merged_cases_5$time_to_event_f, event = rsv_merged_cases_5$rsv_flag)

# Fit Kaplan-Meier model
km_fit <- survfit(surv_obj ~ 1, data = rsv_merged_cases_5)

# Plot KM curve
ggsurvplot(km_fit, 
           data = rsv_merged_cases_5, 
           xlab = "Time (Day)", 
           ylab = "Survival Probability", 
           title = "Kaplan-Meier Curve for RSV Patients", 
           conf.int = TRUE,
           risk.table = TRUE,
           xlim = c(0, 730),   
           ylim = c(0.92, 1), 
           break.y.by = 0.01  
)
```


```{r}
# Fit the KM model stratified by Sex
km_fit_sex <- survfit(surv_obj ~ sex, data = rsv_merged_cases_5)
# Plot the KM curve stratified by Sex
ggsurvplot(km_fit_sex,
           data = rsv_merged_cases_5,
           conf.int = TRUE,
           risk.table = TRUE,
           xlab = "Time (Day)",
           ylab = "Survival Probability",
           title = "Kaplan-Meier Curve Stratified by Sex",
           xlim = c(0, 730),   
           ylim = c(0.92, 1), 
           break.y.by = 0.01  
)
```

```{r}
rsv_merged_cases_5 <- rsv_merged_cases_5 %>%
  mutate(Race = factor(Race, 
                       levels = c("White", "Black", "Asian or Pacific Islander", 
                                  "Hispanic", "Native American", "Other"),
                       ordered = TRUE))

# Fit the KM model stratified by Race
km_fit_race <- survfit(surv_obj ~ Race, data = rsv_merged_cases_5)
# Plot the KM curve stratified by Race
ggsurvplot(
  km_fit_race,
  data = rsv_merged_cases_5,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Day)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by Race",
  xlim = c(0, 730),   
  ylim = c(0.92, 1), 
  break.y.by = 0.01  
)
```

```{r}
# Recode National Income Quartile as an ordered factor
rsv_merged_cases_5 <- rsv_merged_cases_5 %>%
  mutate(
    StateIncomeQuartile = factor(
      StateIncomeQuartile,
      levels = c("$1 – $47,999", "$48,000 – $60,999", "$61,000 – $81,999", "$82,000 and above","Missing"),
      ordered = TRUE
    )
  )

# Fit the KM model stratified by National Income Quartile
km_fit_income <- survfit(surv_obj ~ StateIncomeQuartile, data = rsv_merged_cases_5)

# Plot the KM curve with custom legend labels
ggsurvplot(
  km_fit_income,
  data = rsv_merged_cases_5,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Day)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by National Income Quartile",
  legend.labs = c(
    "Income = $1 – $47,999",
    "Income = $48,000 – $60,999",
    "Income = $61,000 – $81,999",
    "Income = $82,000 and above",
    "Income = Missing"
  ),
  xlim = c(0, 730),
  ylim = c(0.92, 1),
  break.y.by = 0.01
)

```

```{r}
rsv_merged_cases_5 <- rsv_merged_cases_5 %>%
  mutate(Birthweight = factor(Birthweight, 
                       levels = c("Normal", "Low Birthweight", "Very Low Birthweight","Unknown"),
                       ordered = TRUE))

# Fit the KM model stratified by Sex
km_fit_birthweight <- survfit(surv_obj ~ Birthweight, data = rsv_merged_cases_5)

# Plot the KM curve stratified by Sex
ggsurvplot(
  km_fit_birthweight,
  data = rsv_merged_cases_5,
  conf.int = TRUE,
  risk.table = TRUE,
  xlab = "Time (Day)",
  ylab = "Survival Probability",
  title = "Kaplan-Meier Curve Stratified by Birth Weight",
  xlim = c(0, 730),   
  ylim = c(0.92, 1), 
  break.y.by = 0.01  
)
```

