---
title: "Table1"
output: html_document
date: "2025-04-19"
---

saveRDS(SID_rsv_filtered, '/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/SID_rsv_filtered_20250412')
SID_rsv_filtered <- readRDS("/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/SID_rsv_filtered_20250412")


# SID only
```{r}
# SID sociodemographic table

library(dplyr)
library(table1)

SID_rsv_filtered_1 <- SID_rsv_filtered

# Recode RSV flag
SID_rsv_filtered_1$rsv_flag <- factor(
  as.numeric(as.character(SID_rsv_filtered_1$rsv_flag)),
  levels = c(0, 1),
  labels = c("RSV negative", "RSV positive")
)

# Recode sex variable (0 = Male, 1 = Female)
SID_rsv_filtered_1 <- SID_rsv_filtered_1 %>%
  mutate(Sex = factor(sex, levels = c(0, 1), labels = c("Male", "Female")))

# Recode race
SID_rsv_filtered_1 <- SID_rsv_filtered_1 %>%
  mutate(Race = case_when(
    race == "1" ~ "White",
    race == "2" ~ "Black",
    race == "3" ~ "Hispanic",
    race == "4" ~ "Asian or Pacific Islander",
    race == "5" ~ "Native American",
    race == "6" ~ "Other",
    TRUE ~ "Missing"
 )) %>%
  mutate(Race = factor(Race, levels = c(
    "White", 
    "Black", 
    "Hispanic", 
    "Asian or Pacific Islander", 
    "Native American", 
    "Other"
  )))

# Recode income quartile
SID_rsv_filtered_1 <- SID_rsv_filtered_1 %>%
  mutate(StateIncomeQuartile = case_when(
    income_national_quartile == 1 ~ "$1 – $47,999",
    income_national_quartile == 2 ~ "$48,000 – $60,999",
    income_national_quartile == 3 ~ "$61,000 – $81,999",
    income_national_quartile == 4 ~ "$82,000 and above",
    TRUE ~ "Missing"
  )) %>%
  mutate(StateIncomeQuartile = factor(StateIncomeQuartile))

# Birthweight sequence
SID_rsv_filtered_1 <- SID_rsv_filtered_1 %>%
  mutate(Birthweight = factor(Birthweight,
                              levels = c("Normal", "Low Birthweight", "Very Low Birthweight", "Unknown"),
                              ordered = TRUE))

# Apply variable labels
label(SID_rsv_filtered_1$Sex) <- "Sex at Birth"
label(SID_rsv_filtered_1$Race) <- "Race/Ethnicity"
label(SID_rsv_filtered_1$StateIncomeQuartile) <- "State Income Quartile"

# Generate Table: Birthweight as rows, RSV as columns
table1(
  ~ Birthweight + Sex + Race + StateIncomeQuartile | rsv_flag,
  data = SID_rsv_filtered_1,
  overall = "Total",
  caption = "Table 2. Sociodemographics by RSV Status in NY Infants <2Y (SID)"
)

```



saveRDS(SEDD_rsv_filtered, '/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/SEDD_rsv_filtered_')
SEDD_rsv_filtered <- readRDS("/Volumes/home/HCUPdata-CC0941-MEDSPH/myrrha/Data/SEDD_rsv_filtered_")

# SEDD only
```{r}
# SEDD sociodemographic table

library(dplyr)
library(table1)

SEDD_rsv_filtered_1 <- SEDD_rsv_filtered_final

# Recode RSV flag
SEDD_rsv_filtered_1$rsv_flag <- factor(
  as.numeric(as.character(SEDD_rsv_filtered_1$rsv_flag)),
  levels = c(0, 1),
  labels = c("RSV negative", "RSV positive")
)

# Recode sex variable (0 = Male, 1 = Female)
SEDD_rsv_filtered_1 <- SEDD_rsv_filtered_1 %>%
  mutate(Sex = factor(sex, levels = c(0, 1), labels = c("Male", "Female")))

# Recode race
SEDD_rsv_filtered_1 <- SEDD_rsv_filtered_1 %>%
  mutate(Race = case_when(
    race == "1" ~ "White",
    race == "2" ~ "Black",
    race == "3" ~ "Hispanic",
    race == "4" ~ "Asian or Pacific Islander",
    race == "5" ~ "Native American",
    race == "6" ~ "Other",
    TRUE ~ "Missing"
 )) %>%
  mutate(Race = factor(Race, levels = c(
    "White", 
    "Black", 
    "Hispanic", 
    "Asian or Pacific Islander", 
    "Native American", 
    "Other"
  )))

# Recode income quartile
SEDD_rsv_filtered_1 <- SEDD_rsv_filtered_1 %>%
  mutate(StateIncomeQuartile = case_when(
    income_national_quartile == 1 ~ "$1 – $47,999",
    income_national_quartile == 2 ~ "$48,000 – $60,999",
    income_national_quartile == 3 ~ "$61,000 – $81,999",
    income_national_quartile == 4 ~ "$82,000 and above",
    TRUE ~ "Missing"
  )) %>%
  mutate(StateIncomeQuartile = factor(StateIncomeQuartile))

# Birthweight sequence
SEDD_rsv_filtered_1 <- SEDD_rsv_filtered_1 %>%
  mutate(Birthweight = factor(Birthweight,
                              levels = c("Normal", "Low Birthweight", "Very Low Birthweight", "Unknown"),
                              ordered = TRUE))

# Apply variable labels
label(SEDD_rsv_filtered_1$Sex) <- "Sex at Birth"
label(SEDD_rsv_filtered_1$Race) <- "Race/Ethnicity"
label(SEDD_rsv_filtered_1$StateIncomeQuartile) <- "State Income Quartile"

# Generate Table: Birthweight as rows, RSV as columns
table1(
  ~ Birthweight + Sex + Race + StateIncomeQuartile | rsv_flag,
  data = SEDD_rsv_filtered_1,
  overall = "Total",
  caption = "Table 3. Sociodemographics by RSV Status in NY Infants <2Y (SEDD)"
)

```

##### Run and get rsv_merge from PartA first
# SID + SEDD # Alomost no difference to SID
```{r}
library(dplyr)
library(table1)

rsv_merged_1_filtered_1 <- rsv_merged_1_filtered

# Recode RSV flag
# rsv_merged_1_filtered_1$rsv_flag <- factor(
#   as.numeric(as.character(rsv_merged_1_filtered_1$rsv_flag)),
#   levels = c(0, 1),
#   labels = c("RSV negative", "RSV positive")
# )

# # Recode sex variable (0 = Male, 1 = Female)
# rsv_merged_1_filtered_1 <- rsv_merged_1_filtered_1 %>%
#   mutate(sex = factor(sex, levels = c(0, 1), labels = c("Male", "Female")))

# # Recode race using case_when
# rsv_merged_1_filtered_1 <- rsv_merged_1_filtered_1 %>%
#   mutate(Race = case_when(
#     race == "1" ~ "White",
#     race == "2" ~ "Black",
#     race == "3" ~ "Hispanic",
#     race == "4" ~ "Asian or Pacific Islander",
#     race == "5" ~ "Native American",
#     race == "6" ~ "Other",
#     TRUE ~ "Missing"
#   )) %>%
#   mutate(Race = factor(Race))

# # Recode income quartile using case_when
# rsv_merged_1_filtered_1 <- rsv_merged_1_filtered_1 %>%
#   mutate(StateIncomeQuartile = case_when(
#     income_national_quartile == 1 ~ "$1 – $47,999",
#     income_national_quartile == 2 ~ "$48,000 – $60,999",
#     income_national_quartile == 3 ~ "$61,000 – $81,999",
#     income_national_quartile == 4 ~ "$82,000 and above",
#     TRUE ~ "Missing"
#   )) %>%
#   mutate(StateIncomeQuartile = factor(StateIncomeQuartile))

# # Ensure Birthweight is treated as a factor
# rsv_merged_1_filtered_1$Birthweight <- as.factor(rsv_merged_1_filtered_1$Birthweight)
# label(rsv_merged_1_filtered_1$Birthweight) <- "Birthweight"
# 
# # Apply variable labels for display in table1
# label(rsv_merged_1_filtered_1$sex) <- "Sex at Birth"
# label(rsv_merged_1_filtered_1$Race) <- "Race/Ethnicity"
# label(rsv_merged_1_filtered_1$StateIncomeQuartile) <- "State Income Quartile"

# Generate Table 1: Birthweight as row variable, RSV as stratifying variable
table1(
  ~ Birthweight + Sex + Race + StateIncomeQuartile | rsv_flag,
  data = rsv_merged_1_filtered,
  overall = "Total",
  caption = "Table 1. Sociodemographics by RSV Status in NY Infants <2Y"
)

```



====== NY Limited

```{r}
# SID sociodemographic table

library(dplyr)
library(table1)


SID_rsv_filtered_1_NY <- SID_rsv_filtered_1 %>%
  left_join(zip_map, by = c("zip_3digits" = "X3.Digit.ZIP.Code"))

SID_rsv_filtered_1_NY <- SID_rsv_filtered_1_NY%>%
  filter(State.Abbreviation == "NY")

# # Recode RSV flag
# SID_rsv_filtered_1_NY$rsv_flag <- factor(
#   as.numeric(as.character(SID_rsv_filtered_1_NY$rsv_flag)),
#   levels = c(0, 1),
#   labels = c("RSV negative", "RSV positive")
# )

# # Recode sex variable (0 = Male, 1 = Female)
# SID_rsv_filtered_1_NY <- SID_rsv_filtered_1_NY %>%
#   mutate(sex = factor(sex, levels = c(0, 1), labels = c("Male", "Female")))

# # Recode race
# SID_rsv_filtered_1_NY <- SID_rsv_filtered_1_NY %>%
#   mutate(Race = case_when(
#     race == "1" ~ "White",
#     race == "2" ~ "Black",
#     race == "3" ~ "Hispanic",
#     race == "4" ~ "Asian or Pacific Islander",
#     race == "5" ~ "Native American",
#     race == "6" ~ "Other",
#     TRUE ~ "Missing"
#  )) %>%
#   mutate(Race = factor(Race, levels = c(
#     "White", 
#     "Black", 
#     "Hispanic", 
#     "Asian or Pacific Islander", 
#     "Native American", 
#     "Other"
#   )))

# # Recode income quartile
# SID_rsv_filtered_1_NY <- SID_rsv_filtered_1_NY %>%
#   mutate(StateIncomeQuartile = case_when(
#     income_national_quartile == 1 ~ "$1 – $47,999",
#     income_national_quartile == 2 ~ "$48,000 – $60,999",
#     income_national_quartile == 3 ~ "$61,000 – $81,999",
#     income_national_quartile == 4 ~ "$82,000 and above",
#     TRUE ~ "Missing"
#   )) %>%
#   mutate(StateIncomeQuartile = factor(StateIncomeQuartile))

# Birthweight sequence
SID_rsv_filtered_1_NY <- SID_rsv_filtered_1_NY %>%
  mutate(Birthweight = factor(Birthweight,
                              levels = c("Normal", "Low Birthweight", "Very Low Birthweight", "Unknown"),
                              ordered = TRUE))

# Apply variable labels
label(SID_rsv_filtered_1_NY$sex) <- "Sex at Birth"
label(SID_rsv_filtered_1_NY$Race) <- "Race/Ethnicity"
label(SID_rsv_filtered_1_NY$StateIncomeQuartile) <- "State Income Quartile"

# Generate Table: Birthweight as rows, RSV as columns
table1(
  ~ Birthweight + sex + Race + StateIncomeQuartile | rsv_flag,
  data = SID_rsv_filtered_1_NY,
  overall = "Total",
  caption = "Sl Table 2. Sociodemographics by RSV Status in NY Infants <2Y (SEDD, NY Residents)"
)

```


```{r}
library(dplyr)

SID_states <- SID_rsv_filtered_1_NY %>%
  select(id, State.Abbreviation) %>%
  rename(State_SID = State.Abbreviation)

SEDD_rsv_filtered_1_NY <- SEDD_rsv_filtered_1 %>%
  left_join(SID_states, by = "id")

SEDD_rsv_filtered_1_NY <- SEDD_rsv_filtered_1_NY%>%
  filter(State_SID == "NY")

```

```{r}
# SEDD sociodemographic table

library(dplyr)
library(table1)


# # Recode RSV flag
# SEDD_rsv_filtered_1_NY$rsv_flag <- factor(
#   as.numeric(as.character(SEDD_rsv_filtered_1_NY$rsv_flag)),
#   levels = c(0, 1),
#   labels = c("RSV negative", "RSV positive")
# )

# # Recode sex variable (0 = Male, 1 = Female)
# SEDD_rsv_filtered_1_NY <- SEDD_rsv_filtered_1_NY %>%
#   mutate(sex = factor(sex, levels = c(0, 1), labels = c("Male", "Female")))

# # Recode race
# SEDD_rsv_filtered_1_NY <- SEDD_rsv_filtered_1_NY %>%
#   mutate(Race = case_when(
#     race == "1" ~ "White",
#     race == "2" ~ "Black",
#     race == "3" ~ "Hispanic",
#     race == "4" ~ "Asian or Pacific Islander",
#     race == "5" ~ "Native American",
#     race == "6" ~ "Other",
#     TRUE ~ "Missing"
#  )) %>%
#   mutate(Race = factor(Race, levels = c(
#     "White", 
#     "Black", 
#     "Hispanic", 
#     "Asian or Pacific Islander", 
#     "Native American", 
#     "Other"
#   )))

# # Recode income quartile
# SEDD_rsv_filtered_1_NY <- SEDD_rsv_filtered_1_NY %>%
#   mutate(StateIncomeQuartile = case_when(
#     income_national_quartile == 1 ~ "$1 – $47,999",
#     income_national_quartile == 2 ~ "$48,000 – $60,999",
#     income_national_quartile == 3 ~ "$61,000 – $81,999",
#     income_national_quartile == 4 ~ "$82,000 and above",
#     TRUE ~ "Missing"
#   )) %>%
#   mutate(StateIncomeQuartile = factor(StateIncomeQuartile))

# Birthweight sequence
SEDD_rsv_filtered_1_NY <- SEDD_rsv_filtered_1_NY %>%
  mutate(Birthweight = factor(Birthweight,
                              levels = c("Normal", "Low Birthweight", "Very Low Birthweight", "Unknown"),
                              ordered = TRUE))

# Apply variable labels
label(SEDD_rsv_filtered_1_NY$sex) <- "Sex at Birth"
label(SEDD_rsv_filtered_1_NY$Race) <- "Race/Ethnicity"
label(SEDD_rsv_filtered_1_NY$StateIncomeQuartile) <- "State Income Quartile"

# Generate Table: Birthweight as rows, RSV as columns
table1(
  ~ Birthweight + sex + Race + StateIncomeQuartile | rsv_flag,
  data = SEDD_rsv_filtered_1_NY,
  overall = "Total",
  caption = "Table 2. Sociodemographics by RSV Status in NY Infants <2Y (SEDD, NY Residents)"
)

```